AvoidFutRessimpletest

### 2021-12-13 R/AvoidFutRessimpletest.Rmd
Avoid Future Reserve solved

---
title: "AvoidFutRessimpletest"
author: "Vincyane Badouard"
date: "10/12/2021"
output: html_document
---
```{r}
library(Maria)
library(sf)
library(dplyr)
library(nngeo)
library(ggplot2)
```

# Steps before the treefeling function
```{r}
MainTrail <- sf::st_linestring(matrix(c(286400, 582945,
                                        286400, 583250,
                                        286700, 583250,
                                        286700, 582945,
                                        286400, 582945) # the return
                                      ,ncol=2, byrow=TRUE))

pol1 <- list(matrix(c(286503, 583134,
                      286503, 583240,
                      286507, 583240,
                      286507, 583134,
                      286503, 583134) # the return
                    ,ncol=2, byrow=TRUE))
pol2 <- list(matrix(c(286650, 583134,
                      286650, 583240,
                      286654, 583240,
                      286654, 583134,
                      286650, 583134) # the return
                    ,ncol=2, byrow=TRUE))

PolList = list(pol1,pol2) #list of lists of numeric matrices
ScndTrail <- sf::st_multipolygon(PolList)

inventory <- addtreedim(inventorycheckformat(Paracou6_2016),
                        volumeparameters = ForestZoneVolumeParametersTable)

inventory <- suppressMessages(treeselection(inventory, objective = 20, scenario ="manual",
                                            fuel = "2", diversification = TRUE, specieslax = FALSE,
                                            objectivelax = TRUE, topography = DTMParacou, plotslope = PlotSlope,
                                            speciescriteria = SpeciesCriteria,
                                            advancedloggingparameters = loggingparameters())$inventory)

FutureReserveCrowns <- inventory %>% # create an object with future/reserve crowns only
  dplyr::filter(LoggingStatus == "future" | LoggingStatus == "reserve") %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

inventory <- inventory %>%
  dplyr::filter(Selected == "1") %>%
  dplyr::select(idTree,DBH,TrunkHeight,TreeHeight,CrownHeight,
                CrownDiameter,Selected, Xutm, Yutm)

dat <- inventory[2,] %>% # just 1 row (1 tree)
  # force the orientation success for the exemple
  tibble::add_column(TreeFellingOrientationSuccess = "1")
```

Our tree to be felled
```{r}
dat$Xutm <- 286546.1
dat$Yutm <- 583029.1
```
# A buffer around the Foot
```{r}
FootBuffer <- dat %>% 
  mutate(Points = st_point(c(Xutm, Yutm)) %>% 
           st_buffer(dist = CrownDiameter/2) %>% 
           st_cast("MULTIPOINT") %>% 
           st_sfc())

# Multipoints to points
FootBuffer <- st_cast(st_sfc(FootBuffer$Points), "POINT", group_or_split = TRUE)

```

2 av/res trees
```{r}
FutureReserveCrowns <- FutureReserveCrowns[1:2,]

# at the same y than the tree to be felled
# 1 on the right side
FutureReserveCrowns[1,]$Xutm <- dat$Xutm + 30
FutureReserveCrowns[1,]$Yutm <- dat$Yutm + 10

# 1 on the left side
FutureReserveCrowns[2,]$Xutm <- dat$Xutm - 30 
FutureReserveCrowns[2,]$Yutm <- dat$Yutm + 10

@@

```

# Preliminary steps
```{r}
# Function
Trail <- st_union(MainTrail, ScndTrail) # Our trail will be MainTrail or ScndTrail

# The crown
Crown <- dat %>%
  mutate(xCrown = Xutm,
         yCrown = Yutm + TrunkHeight + CrownHeight/2,
         exCrown = CrownDiameter/2,
         eyCrown = CrownHeight/2) %>%
  st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

# The trunk
Trunk <- with(dat,
              st_polygon(list(matrix(c(Xutm-(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm) # the return
                                     ,ncol=2, byrow=TRUE))))

# Find the point (TrailPt) on the Trail closest to the location of the tree (Foot)
Foot <- st_point(c(dat$Xutm,dat$Yutm)) # tree foot point

@@

NearestPoint <- st_cast(NearestPoints, "POINT") # to have start (Foot) and end (TrailPt) points
TrailPt <- NearestPoint[[2]] # the point (TrailPt) on the Trail closest to the location of the tree (Foot)

@@

  theta <- as.numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                    c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))

@@

  theta <- 360 - numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                       c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))
}

## when the tree is on the trail (no angle)
if(is.na(theta)) theta <- 0

```

# Compute points
```{r}
# if (directionalfelling != "0")

@@

# if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
if(any(FutureReserveCrowns$TreeHeight <= dat$TreeHeight)){

  FutResCrownsPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight <= dat$TreeHeight) %>% 
    mutate(Crowns = st_cast(Crowns, "MULTIPOINT")) %>% # polygon to multipoint
    summarise(Points = st_combine(Crowns)) # several multipoints to 1
}

# if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
if(any(FutureReserveCrowns$TreeHeight > dat$TreeHeight)){

  FutResDBHPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight > dat$TreeHeight) %>% 
    rowwise() %>% 
    mutate(Points = st_point(c(Xutm, Yutm)) %>% 
             st_buffer(dist = DBH/2) %>% 
             st_cast("MULTIPOINT") %>% 
             st_sfc()) %>% 
    ungroup() %>% 
    summarise(Points = st_combine(Points)) # several multipoints to 1
}

@@

```

@@

# Distances between each of these points and Foot
FutResDists <- rep(NA, length(FutResAllPoints)) # empty list for distances
RotationAngle <- rep(NA, length(FutResDists)) # empty list for rotation angles
OppAngToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
TrailAnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
AnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid

for (FRP in 1:length(FutResAllPoints)) { # for each point
  FutResDists[[FRP]] <- st_distance(Foot, FutResAllPoints[[FRP]]) # compute the distance for each point

  # If distances < TreeHeight of the tree to be felled
  if(FutResDists[[FRP]] < dat$TreeHeight){
    # sum(FutResDists < dat$TreeHeight) # check av/res trees close to the tree to be felled

@@

        RotationAngle[[FRP]] <- as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], dat$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - FootBuffer[[P]][2]), # position to avoid
          degree = TRUE))

@@

        RotationAngle[[FRP]] <- 360 - as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], dat$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - FootBuffer[[P]][2]), # position to avoid
          degree = TRUE))
      }
    }

    # sum(!is.na(RotationAngle)) # check for how many av/res trees the angle has been computed

    ## Then the angle to the trail: the angle to avoid
    OppAngToAvoid[[FRP]] <- RotationAngle[[FRP]] - 180 - theta # valeurs négatives c'est là le prblm
    TrailAnglesToAvoid[[FRP]] <- 180-(90 + OppAngToAvoid[[FRP]])

    TrailAnglesToAvoid[[FRP]] <- round(TrailAnglesToAvoid[[FRP]], digits = 0)

    if(TrailAnglesToAvoid[[FRP]] >= 360) TrailAnglesToAvoid[[FRP]] <- TrailAnglesToAvoid[[FRP]] - 360

    # Record their angles if there are in the fall zone ([1; 179°])
    if(TrailAnglesToAvoid[[FRP]] > 1 & TrailAnglesToAvoid[[FRP]] < 179){
      # any(!is.na(TrailAnglesToAvoid) < 179)

      AnglesToAvoid[[FRP]] <- TrailAnglesToAvoid[[FRP]]
      # Plus1 <- AnglesToAvoid + 1
      # Moins1 <- AnglesToAvoid - 1
      # Plus2 <- AnglesToAvoid + 2
      # Moins2 <- AnglesToAvoid - 2 # to add a 2° buffer
      # AnglesToAvoid <- c(AnglesToAvoid, Plus1, Moins1) #, Plus2, Moins2
      # sum(!is.na(Plus1)) to check

    }else{AnglesToAvoid <- NULL} # No angles to avoid

    # sum(!is.na(AnglesToAvoid)) # check for how many av/res trees the angle has been computed
  }else{AnglesToAvoid <- NULL} # No angles to avoid
}

```

# Tree felling
```{r}
# Drop the tree between [1; 179°] avoiding the angles of the futures/reserves
if(dat$TreeFellingOrientationSuccess == "1"){

@@

    Aangle <- round(as.numeric(180 + OppAng + theta), digits = 0) # Foot oriented

@@

    Aangle <- round(as.numeric(360 - OppAng + theta), digits = 0) # Foot oriented
  }

@@

# Plot
```{r}
ggplot() +
  geom_sf(data = Foot) +
  geom_sf(data = Trail) +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = FallenTree) +
  geom_sf(data = FutResAllPoints) + # FutureReserveCrowns
  geom_sf(data = FutResCrownsPoints, col = "green") + # smaller trees
  geom_sf(data = FutResDBHPoints, col = "pink") +
  geom_sf(data = FootBuffer)

# geom_sf(data = FutResAllPoints, col = "green")
```

>>>>

---
title: "AvoidFutRessimpletest"
author: "Vincyane Badouard"
date: "10/12/2021"
output: html_document
---
```{r}
rm(list = ls()) ; invisible(gc())
library(Maria)
library(sf)
library(dplyr)
library(nngeo)
library(ggplot2)
set.seed(420)
```

# Trails

```{r}
MainTrail <- sf::st_linestring(matrix(c(286400, 582945,
                                        286400, 583250,
                                        286700, 583250,
                                        286700, 582945,
                                        286400, 582945) # the return
                                      ,ncol=2, byrow=TRUE))
pol1 <- list(matrix(c(286503, 583134,
                      286503, 583240,
                      286507, 583240,
                      286507, 583134,
                      286503, 583134) # the return
                    ,ncol=2, byrow=TRUE))
pol2 <- list(matrix(c(286650, 583134,
                      286650, 583240,
                      286654, 583240,
                      286654, 583134,
                      286650, 583134) # the return
                    ,ncol=2, byrow=TRUE))

PolList = list(pol1,pol2) #list of lists of numeric matrices
ScndTrail <- sf::st_multipolygon(PolList)
Trail <- st_union(MainTrail, ScndTrail) # Our trail will be MainTrail or ScndTrail
rm(MainTrail, ScndTrail, pol1, pol2, PolList)
```

```{r}
ggplot() +
  geom_sf(data = Trail)
```

# Individuals

```{r}
inventory <- addtreedim(inventorycheckformat(Paracou6_2016),
                        volumeparameters = ForestZoneVolumeParametersTable)

inventory <- suppressMessages(treeselection(inventory, objective = 20, scenario ="manual",
                                            fuel = "2", diversification = TRUE, specieslax = FALSE,
                                            objectivelax = TRUE, topography = DTMParacou, 
                                            plotslope = PlotSlope,
                                            speciescriteria = SpeciesCriteria,
                                            advancedloggingparameters = loggingparameters())$inventory)
ind <- inventory %>%
  dplyr::filter(Selected == "1") %>%
  dplyr::select(idTree,DBH,TrunkHeight,TreeHeight,CrownHeight,
                CrownDiameter,Selected, Xutm, Yutm) %>% 
  sample_n(1) %>% 
  tibble::add_column(TreeFellingOrientationSuccess = "1")

ind$Xutm <- 286546.1
ind$Yutm <- 583029.1
Foot <- st_point(c(ind$Xutm,ind$Yutm))
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot)
```

# Future Reserve

```{r}
FutureReserveCrowns <- inventory %>% # create an object with future/reserve crowns only
  dplyr::filter(LoggingStatus == "future" | LoggingStatus == "reserve") %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

FutureReserveCrowns <- FutureReserveCrowns[1:2,]

# at the same y than the tree to be felled
# 1 on the right side
FutureReserveCrowns[1,]$Xutm <- ind$Xutm + 30
FutureReserveCrowns[1,]$Yutm <- ind$Yutm + 10

# 1 on the left side
FutureReserveCrowns[2,]$Xutm <- ind$Xutm - 30 
FutureReserveCrowns[2,]$Yutm <- ind$Yutm + 10

@@

rm(inventory)
```

```{r}
ggplot() +
   geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green")
```

# Individual geometry

```{r}
# The crown
Crown <- ind %>%
  mutate(xCrown = Xutm,
         yCrown = Yutm + TrunkHeight + CrownHeight/2,
         exCrown = CrownDiameter/2,
         eyCrown = CrownHeight/2) %>%
  st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

# The trunk
Trunk <- with(ind,
              st_polygon(list(matrix(c(Xutm-(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm) # the return
                                     ,ncol=2, byrow=TRUE))))

# Find the point (TrailPt) on the Trail closest to the location of the tree (Foot)
 # tree foot point

@@

NearestPoint <- st_cast(NearestPoints, "POINT") # to have start (Foot) and end (TrailPt) points
TrailPt <- NearestPoint[[2]] # the point (TrailPt) on the Trail closest to the location of the tree (Foot)
rm(NearestPoint)

@@

  theta <- as.numeric(matlib::angle(c(Foot[1] - Foot[1], ind$TreeHeight),
                                    c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))

@@

  theta <- 360 - numeric(matlib::angle(c(Foot[1] - Foot[1], ind$TreeHeight),
                                       c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))
}

## when the tree is on the trail (no angle)
if(is.na(theta)) theta <- 0
thetaIndNearest <- theta
rm(theta)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk)
```

# Future Reserve Crowns to Points

```{r}
# if (directionalfelling != "0")

@@

# if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
if(any(FutureReserveCrowns$TreeHeight <= ind$TreeHeight)){

  FutResCrownsPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight <= ind$TreeHeight) %>% 
    mutate(Crowns = st_cast(Crowns, "MULTIPOINT")) %>% # polygon to multipoint
    summarise(Points = st_combine(Crowns)) # several multipoints to 1
}

# if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
if(any(FutureReserveCrowns$TreeHeight > ind$TreeHeight)){

  FutResDBHPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight > ind$TreeHeight) %>% 
    rowwise() %>% 
    mutate(Points = st_point(c(Xutm, Yutm)) %>% 
             st_buffer(dist = DBH/2) %>% 
             st_cast("MULTIPOINT") %>% 
             st_sfc()) %>% 
    ungroup() %>% 
    summarise(Points = st_combine(Points)) # several multipoints to 1
}

@@

rm(FutResCrownsPoints, FutResDBHPoints)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green")
```

# Individual Foot Buffer

```{r}
FootBuffer <- ind %>% 
  mutate(Points = st_point(c(Xutm, Yutm)) %>% 
           st_buffer(dist = CrownDiameter/2) %>% 
           st_cast("MULTIPOINT") %>% 
           st_sfc())
# Multipoints to points
FootBuffer <- st_cast(st_sfc(FootBuffer$Points), "POINT", group_or_split = TRUE)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer)
```

@@

# Distances between each of these points and Foot
RotationAngle <- c() # empty list for rotation angles
for (FRP in 1:length(FutResAllPoints)) { # for each point
  FutResDist <- st_distance(Foot, FutResAllPoints[[FRP]]) # compute the distance for each point

  # If distances < TreeHeight of the tree to be felled
  if(FutResDist < ind$TreeHeight){
    # sum(FutResDists < ind$TreeHeight) # check av/res trees close to the tree to be felled

@@

        RotationAngle <- c(RotationAngle, as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], ind$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - FootBuffer[[P]][2]), # position to avoid
          degree = TRUE)))

@@

        RotationAngle <- c(RotationAngle,  360 - as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], ind$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - 
              FootBuffer[[P]][2]), # position to avoid
          degree = TRUE)))
      }
    }
  }
}
rm(FutResDist)
      # sum(!is.na(RotationAngle)) # check for how many av/res trees the angle has been computed

    ## Then the angle to the trail: the angle to avoid
OppAngToAvoid <- RotationAngle - 180 - thetaIndNearest # valeurs négatives c'est là le prblm
TrailAnglesToAvoid <- 180-(90 + OppAngToAvoid)
TrailAnglesToAvoid <- round(TrailAnglesToAvoid, digits = 0)
TrailAnglesToAvoid[TrailAnglesToAvoid >= 360] <- TrailAnglesToAvoid[TrailAnglesToAvoid >= 360] - 360
AnglesToAvoid <- TrailAnglesToAvoid[
  TrailAnglesToAvoid >= 1 & TrailAnglesToAvoid <= 179
]
```

# One tree felling

```{r}
# Drop the tree between [1; 179°] avoiding the angles of the futures/reserves
if(ind$TreeFellingOrientationSuccess == "1"){

@@

    Aangle <- round(as.numeric(180 + OppAng + thetaIndNearest), digits = 0) # Foot oriented

@@

    Aangle <- round(as.numeric(360 - OppAng + thetaIndNearest), digits = 0) # Foot oriented
  }

@@

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer) +
  geom_sf(data = FallenTree, col = "red")
```

# All tree felling

```{r}
TreefallOrientation <- c(1:179) # the a priori fall zone (angles to the trail)
TreefallOrientation <- TreefallOrientation[!TreefallOrientation %in% unique(AnglesToAvoid)] # Remove angles (the fut/res trees) to avoid

felling <- function(theta){

   OppAng <- 180-(90 + as.numeric(theta)) # the angle between the closest position to the trail (90°) and the desired position (desired angle to the trail)

    Aangle <- round(as.numeric(180 + OppAng + thetaIndNearest), digits = 0) # Foot oriented

    Aangle <- round(as.numeric(360 - OppAng + thetaIndNearest), digits = 0) # Foot oriented
  }

  # Calculate the crown and trunk position
  FallenTree <- st_difference(st_union( # A configuration (90° to the trail (shortest))
    rotatepolygon(Trunk, angle = Aangle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = Aangle, fixed = Foot) # turned crown
  ))
  return(FallenTree)
}

FallenTrees <- lapply(as.list(TreefallOrientation), felling)
FallenTrees <- lapply(FallenTrees, function(x) data.frame(Tree = x)) %>% 
  bind_rows()
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer) +
  geom_sf(data = FallenTrees$geometry, col = "red")
```
